Contention Management for 5G

### 2017-02-01 article/article.tex

	We consider a symmetric directed graph $G=(V,A)$ modelling a network. Each arc  $(u,v)$ in $A$ is labeled by an integer $Dl(u,v) \geq 1$ that we call the delay and
	which represents the number of time slots taken by a signal to go from $u$ to $v$ using this arc. Note that for any arc $(u,v)$, $Dl(u,v)=Dl(v,u)$.

@@

@@

      This affectation represents the following process: first a message is sent in $l$, through the route $r_l$, at time $m_l$.
      This message is received by $\rho(l)$ at time $t(\rho(l),r_l)$. It is then sent back to $l$ at time $\bar{m_l}$,
      which is in the same period if $m_{\rho(l)} > t(\rho(l),r_l)$ otherwise it is in the next. The time between the arrival of
      the message and the time it is sent back is called the \textbf{waiting time} and is defined by $w_l = m_{\rho(l)} - t(\rho(l),r_l)$ if $m_{\rho(l)} > t(\rho(l),r_l)$ and $w_l = m_{\rho(l)} + P - t(\rho(l),r_l)$ otherwise.
      When a BBU receives a message, it must compute the answer before sending it back to the BBU. This time is taken into account
      in the last arc leading to the BBU and thus we need not to consider it explicitely in our model.

@@

   \todo{Definition of the load. La preuve se réfère à la précédente qui n'a pas été retenue, il faut du coup tout expliquer.}

    \begin{theorem}
    Problem PRA cannot be approximated within a factor $n^{1-o(1)}$ unless $\P = \NP$ even when the load is two
    and $n$ is the number of pairs in the assignment.
    \end{theorem}

    \begin{proof}
    We reduce PRA to graph coloring. Let $G$ be a graph instance of the $k$-coloring problem. 
    We define $H$ in the following way: for each vertex $v$ in $G$, there is a route $r_v$ in $H$.
    Two routes $r_v$ and $r_u$ share an edge if and only if $(u,v)$ is an edge in $G$ and this edge is only in this two routes. 
    We put a weight inbetween shared edges in a route so that there is a delay $k$ between two such edges.

    As in the previous proof, a $k$-coloring of $G$ gives a $k$-periodic schedule of $H$
    and conversly. Therefore if we can approximate the value of PRA  within a factor $f$,
    we could approximate the minimal number of colors needed to color a graph within a fator $f$, 
    by doing the previous reduction for all possible $k$. The proof follows from the hardness of approximability
    of finding a minimal coloring~\cite{zuckerman2006linear}.
    \end{proof}

  \subsection{MIN-PRA}
    Exemple de cas simple

@@

  \subsection{No waiting times}
    \subsubsection{Shortest-longest}
      \paragraph{Algo}
      \paragraph{Period}
    \subsubsection{Exhaustive generation}
      Décrire l'algo, expliquer les coupes
    \subsubsection{Results}
      Resultats des simulations : Shortest-longest optimal pour ces parametres.

>>>>

	We consider a directed graph $G=(V,A)$ modelling a network. Each arc  $(u,v)$ in $A$ is labeled by an integer $Dl(u,v) \geq 1$ that we call the delay and
	which represents the number of time slots taken by a signal to go from $u$ to $v$ using this arc. Note that for any arc $(u,v)$, $Dl(u,v)=Dl(v,u)$.

@@

      \todo{Doit on définir un objet graphe + routes + assignement pour ensuite simplifier l'écriture}

@@

      This affectation represents the following process: first a message is sent in $l$, through the route $r_l$, at time $m_l$.
      This message is received by $\rho(l)$ at time $t(\rho(l),r_l)$. It is then sent back to $l$ in the same period at time $m_{\rho(l)}$ if $m_{\rho(l)} > t(\rho(l),r_l)$, otherwise at time $m_{\rho(l)}$ in the next period. The time between the arrival of the message and the time it is sent back is called the \textbf{waiting time} and is defined by $w_l = m_{\rho(l)} - t(\rho(l),r_l)$ if $m_{\rho(l)} > t(\rho(l),r_l)$ and $w_l = m_{\rho(l)} + P - t(\rho(l),r_l)$ otherwise.
      When a BBU receives a message, it must compute the answer before sending it back to the RRH. This time can be encoded
      in the last arc leading to the BBU and thus we need not to consider it explicitely in our model.

@@

 In this section we assume that the size of a message $\tau$ is equal to one, which implies hardness of PRA and PALL for all $\tau$. 
Consider an instance of Problem PRA, i.e., a digraph $G=(V,A)$, an assignment $\cal C$, a routing function $\cal R$ and a period $P$. 
The {\bf conflict depth} of a route is the number of other routes which share an edge with it. 
The conflict depth of an assignment $\cal C$ is the maximum of the conflict depth of the routes in ${\cal R}_{\cal C}$.
The {\bf load} of an assignment is the maximal number of routes sharing the same arc.
It is clear that a $P$-periodic affectation must satisfy that $P$ is larger or equal to the load.

We give two alternate proofs that PRA is $\NP$-complete.
The first one works for conflict depth $2$ and is minimal in this regards since we later prove that for conflict depth one,
it is easy to solve PRA. The second one reduces the problem to graph coloring and implies inapproximability when one tries to minimize the parameter $P$. \\

 \begin{proposition}
Problem PRA is $\NP$-complete, when the routing is of conflict depth two.
\end{proposition}
 \begin{proof}
 The problem $PRA$ is in $\NP$ since given an offfset for each route in an affectation, it is easy to check in linear time in the number of edges whether there are collisions.

  Let $H=(V,E)$ be a graph and let $d$ be its maximal degree. We consider the problem to determine whether $H$ is edge-colorable
  with $d$ or $d+1$ colors. The edge coloring problem is $\NP$-hard~\cite{holyer1981np} and we reduce it to PRA to prove its $\NP$-hardness. We define from $H$ an instance of PRA as follows. The graph $G$ has for vertices $V'= \{ v_1, v_2 \mid v \in V  \} \cup \{ l_{u,v}, s_{u,v} \mid (u,v) \in E \}$ that is two vertices for each vertex and for each edge of $H$. 
  Let $A$ bet the set of arcs of $G$, defined by 
  $$A = \{(v_1,v_2) \mid v\in V\} \cup \{(u_2,v_1)\mid u \neq v \in V^2\} \cup \{(l_{u,v},u_1),(v_2,s_{u,v}) \mid (u,v) \in E \}. $$
  All these arcs are of weigth $0$. 
  For each edge $(u,v) \in E$, the is a route $r_{u,v} = s_{u,v},u_1,u_2,v_1,v_2,l_{u,v}$ in ${\cal R}$.  
  The affectation ${\cal C}$ is the set of pair of vertices $(s_{u,v}, l_{u,v})$.

  Observe that the existence of a $d$-coloring of $H$ is equivalent to the existence of a $d$-periodic affectation
  for $(G,{\cal R},{\cal C})$. Indeed, a $d$-coloring of $H$ can be seen as a labeling of its edges by the integers
  in $\{0,\dots,d-1\}$ and we have a correspondance between a $d$-coloring of $H$ and offests for the routes of $(G,{\cal R},{\cal C})$.
  By contruction, the constraint of having no collision between the routes is equivalent to the fact that no two adjacent edges have
  the same color. Therefore we have reduced edge coloring to PRA which concludes the proof. 
 \end{proof}
 \todo{Faire un dessin d'illustration ?}

 Remark that we have used zero weigth in the proof. If the weigths must be strictly positive, which makes more sense in our model since
 we model the latency of physical links, it is easy to adapt the proof. We just have to set them so that in any route the delay at $u_1$ is equal to $d$ and thus equal to $0$ modulo $d$. We now lift this hardness result to the problem PALL.

\begin{corollary}
Problem PALL is $\NP$-complete for routing of conflict depth two.
\end{corollary}
\begin{proof}
 We consider $(G,{\cal R},{\cal C},P)$ an instance of $PRA$ such that no element appears both in the first and second position in a pair of ${\cal C}$. Remark that this condition is satisfied in the previous proof, which makes the problem $PRA$ restricted to these instances $\NP$-complete. 
 Let us define $T_{max} = 2 \times \max_{r \in {\cal R}} \lambda(r) + P$. We define $\rho$ as the function which maps 
 $u$ to $v$ when $(u,v) \in {\cal C}$. The instance $(G,{\cal R},\rho,P,T_{max})$ is in PALL if and only if $(G,{\cal R},{\cal C},P)$
 is in $PRA$. Indeed the waiting time of each route is by definition less than $P$ and thus the maximal process time less than $T_{max}$. Therefore the fact that $(G,{\cal R},\rho,P,T_{max})$ is in PALL is equivalent to the existence of a $P$-periodic assignment of ${\cal C}_{\rho}$ which is equal to ${\cal C}$.
\end{proof}
% 
% 
% As another corollary of the previous proposition,  given $G$, $\cal R$, $P$ and two disjoint subsets $S$ and $L$ in $G$, the problem of  knowing if there is a matching $\rho$  from $L$ in $S$ such that the answer to problem PRA for instance $G,{\cal C_{\rho}},{\cal R}, P$ is  $\NP$-hard since checking a potential feasible solution for this problem implies to solve PRA.\\

Let MIN-PRA be the problem, given a graph, a routing and an affectation, to find the minimal period $P$ such that there is a $P$-periodic affectation.

\begin{theorem}
 The problem MIN-PRA cannot be approximated in polynomial time within a factor $n^{1-o(1)}$, with $n$ the number of routes, unless $\P = \NP$ even when the load is two.
\end{theorem}

\begin{proof}
 We reduce graph coloring to PRA. Let $H$ be a graph instance of the $k$-coloring problem. 
 We define $G$ in the following way: for each vertex $v$ in $H$, there is a route $r_v$ in $G$.
 Two routes $r_v$ and $r_u$ share an arc if and only if $(u,v)$ is an edge in $H$; this arc is the only one shared by these two routes.   
 All arcs are of delay $0$.

 Observe that the existence of a $k$-coloring of $H$ is equivalent to the existence of a $k$-periodic affectation in $G$, 
 by converting an offset of a route into a color of a vertex and reciprocally. Therefore if we can approximate the minimum value of $P$ within a factor $f$,
 we could approximate the minimal number of colors needed to color a graph within a factor $f$, by doing the previous reduction for all possible $k$. The proof follows from the hardness of approximability of finding a minimal coloring~\cite{zuckerman2006linear}.
\end{proof}

In particular, this reduction shows that even with small maximal load, the 
minimal period can be large.

  \subsection{MIN-PRA}
    Exemple de cas polynomiaux

@@

  \subsection{No waiting times}
    \subsubsection{Shortest-longest}
      \paragraph{Algo}
      \paragraph{Period}
    \subsubsection{Greedy Algorithm with higher bound}
      \paragraph{Algo}
	Todo
      \paragraph{Period}
	This algorithm gives us a solution without waiting times in a maximum period $3.\tau.N$, if we have N routes.\\
	Suppose that we have a period of $3.\tau.N$ slots, divided in $\tau$ macro-slots.
    \subsubsection{Exhaustive generation}
      Décrire l'algo, expliquer les coupes
    \subsubsection{Results}
      Resultats des simulations : Shortest-longest optimal pour ces parametres.
